### 2015-04-18 implementation-packages.tex
edit 3.1
props to homebrew

In Spack, packages are Python scripts that build software artifacts.  
Each package is a class that extends a generic {\tt Package}
base class.  {\tt Package} implements the bulk of the build process, but
subclasses provide their own {\tt install} method to handle the
specifics of particular packages. The subclass does {\it not} have to
handle managing the install location.  Rather, Spack passes the {\tt install}
method a {\tt prefix} parameter.  The package implementor must ensure that
her {\tt install} function installs the package {\it into} the {\tt prefix},
but Spack ensures that {\tt prefix} is computed in such a way that it is
unique for every configuration of a package.  To further simplify 
package implementation, Spack implements an embedded domain-specific
language (DSL).  The DSL provides special directives such as {\tt depends\_on},
{\tt version}, {\tt provides}, and {\tt patch} that add metadata
to the package class.

Figure~\ref{fig:mpileaks} shows the package for \mpileaks, a tool developed
at LLNL to find memory leaks in MPI programs.
The {\tt MpiLeaks} class first provides simple metadata: a text description,
a homepage, and a download URL.  
Next, two {\tt version} directives identify known versions and provide
MD5 checksums used to verify downloads.
Below this, two {\tt depends\_on}
directives indicate prerequisite packages that must be installed before \mpileaks.
Last, the {\tt install()} method contains the commands used to build
the tool.  Spack's DSL allows shell commands to be invoked as Python functions,
and the {\tt install()} method invokes {\tt configure},
{\tt make}, and {\tt make install} commands as a shell script would.

>>>>

In Spack, packages are Python scripts that build software artifacts.  
Each package is a class that extends a generic {\tt Package}
base class.  {\tt Package} implements the bulk of the build process, but
subclasses provide their own {\tt install} method to handle the
specifics of particular packages. The subclass does {\it not} have to
handle managing the install location.  Rather, Spack passes the {\tt install}
method a {\tt prefix} parameter.  The package implementor must ensure that
her {\tt install} function installs the package {\it into} the {\tt prefix},
but Spack ensures that {\tt prefix} is computed in such a way that it is
unique for every configuration of a package.  To further simplify 
package implementation, Spack implements an embedded domain-specific
language (DSL), similar to that of Homebrew~\cite{homebrew}.
The DSL provides special directives such as {\tt depends\_on},
{\tt version}, and {\tt provides} that add metadata
to the package class.

Figure~\ref{fig:mpileaks} shows the package for \mpileaks, a tool developed
at LLNL to find leaks of MPI objects in programs.
The {\tt MpiLeaks} class first provides simple metadata: a text description,
a homepage, and a download URL.  
Next, two {\tt version} directives identify known versions and provide
MD5 checksums used to verify downloads.
Below this, two {\tt depends\_on}
directives indicate prerequisite packages that must be installed before \mpileaks.
Last, the {\tt install()} method contains the commands used to build
the tool.  Spack's DSL allows shell commands to be invoked as Python functions,
and the {\tt install()} method invokes {\tt configure},
{\tt make}, and {\tt make install} commands as a shell script would.
